Distributed multi-objective evolutionary optimization using island-based selective operator application

### 2017-11-27 hpmoon-asoco.tex
Addressing @JJ 's comments on cover letter

One of the most extended
methods to deal with high-dimensional multi-objective optimization problems are the distributed co-evolutionary algorithms. In these methods each
processor deals only with a part of the decision space, that is, the
individuals belonging to the same sub-population (island) explore the
same subset of decision variables. However, only decomposable problems could be strictly implemented in parallel this way. %these methods only works with decomposable problems.

A new approach is presented here to deal with un-decomposable problems, the selective operator application:
instead of sharing parts of individuals, each processor applies the variation 
operators to a specific subset of the whole individual. We have previously proved 
that extending the length of these subsets, and therefore, sharing some search 
space between islands, implies an improvement in the results.

Specifically, in this paper, a new method to
automatically adapt the size of the overlapping, or shared sections,
of the individuals has been compared to other collaborative
evolutionary techniques. Different number of islands, chromosome sizes
and problems have been used. The analysis of the obtained experimental
results, by using different metrics, shows that our approach can provide statistically significant improvements with
respect to the base algorithm.%, being the method to automatically adapt
%the overlapping size the one to obtain the best performance during the
%same execution time. 
 Results also show that the relation between the number
of islands (subpopulations) and the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables. 
\end{abstract}

@@

A similar approach used to solve this kind of problems is the {\em selective operator application} (SLA). In this case, each island deals with the whole chromosome, but only modifies a fragment of it in the crossover and mutation phase depending on the number of islands, using the whole chromosome for fitness calculation. This allows to deal with un-decomposable problems.
In our previous work \citep{Garcia16hpmoon} we used this method in a preliminary experimental setup, although it is worth mentioning we are establishing its name and acronym in this paper for the first time.  We proved
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. We discovered that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, the experiments
were conducted in a limited experimental system: mono-processor and synchronous island model. This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more complete parameter setting. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically adapts the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions. 
% We have to
                                % emphasize what is actually done. Is
                                % this an addition to the old paper,
                                % or a complete rewrite? - JJ FERGU: rewritten previous paragraph

@@

In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two different dMOEAs. The first
one divided the chromosome into $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome (the selective operator application), while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, selective operator application with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivated us to find a new adaptive method to select this number of
sections of the chromosome to use, depending on the number of
islands. Besides, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and on this occasion, the experiments have been run on a parallel
cluster. Therefore, at the same time we are proposing an adaptive
method to divide the individual search space, and also validating the
previous methods.

>>>>

One of the most extended methods to deal with high-dimensional multi-objective optimization problems are the distributed co-evolutionary algorithms. In these methods each processor deals only with a part of the decision space, that is, the individuals belonging to the same sub-population (island) explore the same subset of decision variables. However, only decomposable problems could be strictly implemented in parallel this way. A new approach is presented here to deal with un-decomposable problems, the selective operator application: instead of sharing parts of individuals, each processor applies the variation operators to a specific subset of the whole individual. We have previously proved that extending the length of these subsets, and therefore, sharing some search space between islands, implies an improvement in the results. Specifically, in this paper, a new method to automatically adapt the size of the overlapping, or shared sections, of the individuals has been compared to other collaborative evolutionary techniques. Different number of islands, chromosome sizes and problems have been used. The analysis of the obtained experimental results, by using different metrics, shows that our approach can provide statistically significant improvements with respect to the base algorithm.  Results also show that the relation between the number of islands (subpopulations) and the length of the chromosome (number of decision variables) is also a relevant factor to determine the most efficient alternative to distribute the decision variables. 
\end{abstract}

@@

A similar approach used to solve this kind of problems is the {\em selective operator application} (SLA). In this case, each island deals with the whole chromosome, but only modifies a fragment of it in the crossover and mutation phase depending on the number of islands, using the whole chromosome for fitness calculation. This allows to deal with un-decomposable problems.
In our previous work \citep{Garcia16hpmoon} we used this method in a preliminary experimental setup, although it is worth mentioning we are establishing its name and acronym in this paper for the first time.  We proved
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. We discovered on that paper that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and synchronous island model. This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more complete parameter setting. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically adapts the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions. 
% We have to
                                % emphasize what is actually done. Is
                                % this an addition to the old paper,
                                % or a complete rewrite? - JJ FERGU: rewritten previous paragraph

@@

In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two different dMOEAs. The first
one divided the chromosome into $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome (the selective operator application), while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, selective operator application with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivated us to find a new adaptive method to select this number of
sections of the chromosome to use, depending on the number of
islands. Besides, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and on this occasion, the experiments have been run on a parallel
cluster. Therefore, at the same time we are proposing an adaptive
method to divide the individual search space, and also validating the
previous approach.
